Investigating Overfitting in Convolutional Neural Networks

### 2025-05-23 paper_formatted.tex
added p-value calculation and rough conclusion paragraph.

    We used the following equation to calculate the Standard Error $SE_\beta$ of the slope $\beta$:
    \[
        \mathrm{SE}_{\beta}
        =
        \frac{
            \displaystyle S
        }{
            \displaystyle S_x \sqrt{n - 1}
        }
    \]
    \noindent We need the standard deviation of the residuals and of the parameter counts to find $SE_\beta$, which we found like this:
    \begin{align*}
        \mathrm{S} &=
        \sqrt{
            \frac{
                \sum_{i=1}^n (y_i - \hat{y}_i)^2
            }{
                n - 2
            }
        } \\[1em]
        \mathrm{S}_x &=
        \sqrt{
            \frac{
                \sum_{i=1}^n (x_i - \bar{x})^2
            }{
                n - 1
            }
        }
    \end{align*}
    \noindent After plugging in all of the data from the 150 models, we got $S = 0.01703$ and $S_x = 0.6972$.
    Now we can use those to find $SE_\beta$:
    \[
        \mathrm{SE}_{\beta}
        =
        \frac{
            \displaystyle 0.01703
        }{
            \displaystyle 0.6972 \sqrt{150 - 1}
        }
        =
        0.002008
    \]
    \noindent And now we can find the t-statistic:
    \begin{gather*}
        \mathrm{t} = \frac{b}{SE_\beta} \\[1em]
        \mathrm{t} = \frac{0.008312}{0.002008} = 4.139
    \end{gather*}

@@

>>>>

    \subsubsection*{5.2.1 Standard Error}

    We used the following equation to calculate the Standard Error $SE_\beta$ of the slope $\beta$:
    \[
        \mathrm{SE}_{\beta}
        =
        \frac{
            \displaystyle S
        }{
            \displaystyle S_x \sqrt{n - 1}
        }
    \]
    \noindent We need the standard deviation of the residuals and of the parameter counts to find $SE_\beta$, which we found like this:
    \begin{align*}
        \mathrm{S} &=
        \sqrt{
            \frac{
                \sum_{i=1}^n (y_i - \hat{y}_i)^2
            }{
                n - 2
            }
        } \\[1em]
        \mathrm{S}_x &=
        \sqrt{
            \frac{
                \sum_{i=1}^n (x_i - \bar{x})^2
            }{
                n - 1
            }
        }
    \end{align*}
    \noindent After plugging in all of the data from the 150 models, we got $S = 0.01703$ and $S_x = 0.6972$.
    Now we can use those to find $SE_\beta$:
    \[
        \mathrm{SE}_{\beta}
        =
        \frac{
            \displaystyle 0.01703
        }{
            \displaystyle 0.6972 \sqrt{150 - 1}
        }
        =
        0.002008
    \]

    \subsubsection*{5.2.2 $t$-statistic}
    \noindent And now we can find the t-statistic, which represents how many standard deviations away our sample slope $b$ is from the population slope $\beta$, assuming that $\beta=0$:
    \begin{gather*}
        \mathrm{t} = \frac{b}{SE_\beta} \\[1em]
        \mathrm{t} = \frac{0.008312}{0.002008} = 4.139
    \end{gather*}

    \subsubsection*{5.2.3 $p$-value}
    \noindent Now that we have the t-statistic, we can use it to find the $p$-value, which represents the probability of getting a test statistic
    of $4.139$ or higher, assuming that the population slope $\beta$ is $0$. Because we have 150 data points, we have $150 - 2 = 148$ degrees of freedom.
    Thus, we calculate the $p$-value by finding the amount of area under a $t$-distribution with 148 degrees of freedom, above $t=4.139$. The graph below demonstrates this:

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.7\textwidth]{Images/tdistribution}
        \caption{p-value calculation on a t-distribution with df = 148}
        \label{fig:t_dist}
    \end{figure}

    \noindent Based on the area under the curve in \textit{\autoref{fig:t_dist}} above $t=4.139$, our $p$-value is $2.916\cdot10^{-5}$.

@@

    The $p$-value of $2.916\cdot10^{-5}$ indicates that, assuming that the population slope $\beta=0$, there is around a
    $0.002916\%$ probability of getting a sample slope as great or greater than our sample slope of $0.008312$, in a sample of
    150 randomly selected models from the population of convolutional models of our structure and of size 1 million parameters to 25 million parameters, trained
    on a randomly selected subset of 100 images from the CIFAR-10 train dataset and tested on the full test dataset of CIFAR-10.
    Because the $p$-value of $2.916\cdot10^{-5}$ is less than $\alpha=0.05$, we reject the null hypothesis, which states that the population slope $\beta$
    of the least-squares regression line that relates parameter count to difference in train and test accuracy (train - test) is $0$,
    in the population defined above.
    The data do provide convincing evidence that the population slope $\beta$ is greater than $0$, and that there is a positive correlation between
    parameter count and difference between train and test accuracy (train - test) in the population defined above.
